R Notebook

### 2020-12-13 AdvendOfCode2020/day09.Rmd
Day 09 | commented code

```{r}

library(purrr)
library(magrittr)

input <- as.double(readLines("./day09_input.txt"))

checkData <- function(.preamble, .number) {
  return(.number %in% colSums(combn(.preamble,2)))
}

pre.size <- 25

numChecked <- (pre.size+1):length(input) %>% 
  map_lgl(function(.i, .d, .ps){
    checkData(.d[(.i-.ps):(.i-1)], .d[.i]) %>% 
      return()
  }, .d=input, .ps=pre.size)

input[min(pre.size + which(numChecked==FALSE))]
```

@@

```{r}

library(purrr)
library(magrittr)

input <- as.double(readLines("./day09_input.txt"))

checkData <- function(.preamble, .number) {
  return(.number %in% colSums(combn(.preamble,2)))
}

pre.size <- 25

numChecked <- (pre.size+1):length(input) %>% 
  map_lgl(function(.i, .d, .ps){
    checkData(.d[(.i-.ps):(.i-1)], .d[.i]) %>% 
      return()
  }, .d=input, .ps=pre.size)

step1 <- input[min(pre.size + which(numChecked==FALSE))]

rangeComb <- combn(1:length(input),2)

rngChecked <- rangeComb %>% 
  apply(2,function(.range, .data, .numCheck){
    return(sum(.data[.range[1]:.range[2]])==.numCheck)
  }, .data=input, .numCheck=step1)

answerRange <- rangeComb[,which(rngChecked==T)]
contRange <- input[answerRange[1]:answerRange[2]]

resp <- min(contRange) + max(contRange)

resp

>>>>

```{r}
library(purrr)
library(magrittr)

# reads the input data as an array of numbers
input <- as.double(readLines("./day09_input.txt"))

# function that checks if a number is the sum
# of all combinations of two from the preamble
checkData <- function(.preamble, .number) {
  return(.number %in% colSums(combn(.preamble,2)))
}

# preamble size
pre.size <- 25

# checks, along the data it there is a sum in the preamble
# numCheck return an array of result check for all data
numChecked <- (pre.size+1):length(input) %>% # the range of the data (input-preamble)
  map_lgl(function(.i, .d, .ps){
    # calc the preamble and the number to be checked
    checkData(.d[(.i-.ps):(.i-1)], .d[.i]) %>% 
      return()
  }, .d=input, .ps=pre.size)

# once checked all number, get the first index that fail and
# returns that input value at this position
input[min(pre.size + which(numChecked==FALSE))]
```

@@

```{r}
library(magrittr)

# reads the input data as an array of numbers
input <- as.double(readLines("./day09_input.txt"))

# function that checks if a number is the sum
# of all combinations of two from the preamble
checkData <- function(.preamble, .number) {
  return(.number %in% colSums(combn(.preamble,2)))
}

# preamble size
pre.size <- 25

# checks, along the data it there is a sum in the preamble
# numCheck return an array of result check for all data
numChecked <- (pre.size+1):length(input) %>% # the range of the data (input-preamble)
  map_lgl(function(.i, .d, .ps){
    # calc the preamble and the number to be checked
    checkData(.d[(.i-.ps):(.i-1)], .d[.i]) %>% 
      return()
  }, .d=input, .ps=pre.size)

# once checked all number, get the first index that fail and
# returns that input value at this position
# store the answer in the 'step1' var to be used in sequence
step1 <- input[min(pre.size + which(numChecked==FALSE))]

# Now we must find a continuous range of falues which the sum is equal
# the value found in step 1

# Calc all possible "continuous" range index
rangeComb <- combn(1:length(input),2)

# for each, check if the sum of its values match the target value
# we test all then and store the test result
rngChecked <- rangeComb %>% 
  apply(2,function(.range, .data, .numCheck){
    return(sum(.data[.range[1]:.range[2]])==.numCheck)
  }, .data=input, .numCheck=step1)

# get the index range which the range matched the sum
answerRange <- rangeComb[,which(rngChecked==T)]

# get the values in this range
contRange <- input[answerRange[1]:answerRange[2]]

# sum the smallest and the largest value in this range
resp <- min(contRange) + max(contRange)

# this is our answer
resp
